---
title: "Homework #2"
author: David Nemirovsky
date: 9/30/20
output: github_document
---

```{r setup, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset:

```{r trashwheel data}
trashwheel_df = 
  read_xlsx(
    "./hw2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls))
```

Read the precipitation data:

```{r precip data}
precip_2018 = 
  read_excel(
    "./hw2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation",
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./hw2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2017 Precipitation",
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now, combine annual precipitation:

```{r combined precip data}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name)

precip_df = 
  bind_rows(precip_2018, precip_2017)

precip_df =
left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector located in Baltimore, Maryland. Mr. Trashwheel collects trash as it enters the inner harbor and stores it in a dumpster. This dataset contains information on year, month, and trash collected, including some specific types of trash. There are a total of `r nrow(trashwheel_df)` rows in our data set. Additional data sheets include monthly precipitation data. The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches. The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.

## Problem 2

Read in, clean names, and select the appropriate NYC Transit dataset:

```{r read and select data, message = FALSE}
transit_df = 
  read_csv("./hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  unite(route, route1:route11, sep = ", ", na.rm = TRUE) %>% 
  relocate(station_name, route) %>% 
  arrange(station_name, route) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

This dataset contains information about the entrances of all of the NYC subway stations. Some key variables are subway lines, routes, types of entry, presence of MetroCard vending machines, and whether or not that station entrance is ADA compliant. A variety of functions were used in order to tidy this data, including `clean_names` from the `janitor` package to rename the variables in snake-case. The functions`select`, `unite`, `relocate`, `arrange`, and `mutate` (from the `dplyr` package in `tidyverse`) were used to choose the appropriate data to study, collapse the "route#" variables into a single variable indicating train routes, re-position and rearrange the columns into a more appealing order, and convert the "entry" variable from a character to a logical one, respectively. The dataset is composed of `r nrow(transit_df)` rows and `r ncol(transit_df)` columns.

#### Questions About Data:

1. There are `r distinct(transit_df, station_name, route) %>% nrow()` distinct stations.
2. Of these, only `r filter(transit_df, ada == TRUE) %>% distinct(station_name, route) %>% nrow()` stations are ADA compliant.
3. Approximately `r round((filter(transit_df, vending == "NO", entry == TRUE) %>% nrow()) / (filter(transit_df, vending == "NO") %>% nrow()),3)*100`% of station entrances/exits without vending allow entrance.

Now, reformat data to create separate variables for "route number" and "route name":

```{r reformat data, message = FALSE}
transit_reformed_df = 
  read_csv("./hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(across(route8:route11, as.character)) %>% 
  pivot_longer(route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name",
    values_drop_na = TRUE) %>% 
  relocate(station_name, route_number, route_name) %>% 
  arrange(station_name, route_number, route_name) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

There are `r filter(transit_reformed_df, route_name == "A") %>% distinct(station_name, route_number) %>% nrow()` distinct stations that serve the A train. Of these, only `r filter(transit_reformed_df, route_name == "A", ada == TRUE) %>% distinct(station_name, route_number) %>% nrow()` stations are ADA compliant.

## Problem 3

Read in and clean the politician dataset:

```{r clean pol data, message =  FALSE}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name)

pol_df = 
  read_csv("./hw2_data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(across(year:day, as.integer)) %>% 
  left_join(month_df, by = "month") %>% 
  relocate(year:day, prez_gop, prez_dem) %>% 
  pivot_longer(
    prez_gop:prez_dem,
    names_to = "president",
    names_prefix = "prez_",
    values_to = "party") %>% 
  mutate(party = na_if(party, 0)) %>% 
  drop_na(party) %>% 
  relocate(year, month_name, president) %>% 
  select(year:rep_dem, -month, -day)
```

