---
title: "Homework #2"
author: David Nemirovsky
date: 9/30/20
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset:

```{r trashwheel data}
trashwheel_df = 
  read_xlsx(
    "./hw2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls))
```

Read the precipitation data:

```{r precip data}
precip_2018 = 
  read_excel(
    "./hw2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation",
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./hw2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2017 Precipitation",
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now, combine annual precipitation:

```{r combined precip data}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name)

precip_df = 
  bind_rows(precip_2018, precip_2017)

precip_df =
left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector located in Baltimore, Maryland. Mr. Trashwheel collects trash as it enters the inner harbor and stores it in a dumpster. This dataset contains information on year, month, and trash collected, including some specific types of trash. There are a total of `r nrow(trashwheel_df)` rows in our data set. Additional data sheets include monthly precipitation data. The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches. The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.